---
title: "Final Group Project Data Mining"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
author: "Adriana Barlafante, Guglielmo Bonifazi, Niklas Rengelshausen"
date: "1/11/2022"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# World Happiness Report

**Task:** The aim is to study the characteristics of society described
in the data sets: what factors affect well-being and how do they relate
to each other? The two bases provided should be combined.

## Introduction

For our analysis of the data set we divided our work into multiple
steps. To have an overview of our steps please have a look at the Index.
The data analysis is made with two data set which we later combined
after having a first look at both data sets and the kaggle page to
better understand the data.

**Origin of the data:** World Happiness (WH):
<https://www.kaggle.com/unsdsn/world-happiness> Human Freedom Index
(HFI): <https://www.kaggle.com/gsutters/the-human-freedom-index>

In addition we had a look at the used the [following
book](https://www.cato.org/sites/cato.org/files/2021-01/human-freedom-index-2016.pdf)
to better understand the data set of the Human Freedom Index. The data
set contains over 120 different variables and to understand these
variables the book helped with some descriptions (see p. 185 ff.)

# Index

-- 1. Filtering of the Data Set and Imputation of Values 
-- 2. Descriptive Statistics 
-- 3. Principal Component Analysis (PCA) 
-- 4. Cluster Analysis 
-- 5. Multidimensional Scaling 
-- 6. Perceptual Maps 
-- 7. Possible Extension (NEW HEADLINE NEEDED)

```{r include=FALSE, warning=FALSE}
## PLEASE LOAD LIBRARY HERE AND RERUN CODE CHUNK
library(data.table)
library(tidyverse)
```

## 1. Filtering of the Data Set and Imputation of Values

### 1.1 Loading Data Sets and First Look

```{r}
wh = read.csv("2016.csv")         # World Happiness
hfi = read.csv("hfi_cc_2018.csv") # Human Freedom Index
```

**World Happiness:**

```{r include=TRUE, echo=FALSE}
wh
```

This data set contains information regarding the happiness of different
countries. Next to an index score which ranks each country the data set
also has a score happiness score on which the happiness rank is based.
Additionally the data set contains columns such as GDP or Health Life
Expectancy. All the observations are from 2016. For a further insights
please have a look at the descriptive statistics.

**Human Freedom Index:**

```{r}
hfi
```
For the second data set we cans see that it contains 1458 observations and 123 variables. The data set contains a 
huge number of variables. Nonetheless, they are organized hierarchically so we do not need them all. As previously
stated, the data set deals with the human freedom index. 

This variable has been computed considering two factors:personal freedom and economic freedom. 
We can observe that human freedom (hf) is the simple average of economic freedom (ec) and personal freedom (pf).

```{r}
all.equal(hfi$hf_score, (hfi$ef_score + hfi$pf_score)/2)
```

More in detail, economic freedom and personal freedom have been computed using other factors.
```{r}
#Indexes of variables used to compute economic freedom
ec=c(70,81,86,99,118)
colnames(hfi)[ec]
```

```{r}
#Indexes of variables used to compute personal freedom
pf=c(8,22,26,32,44,52,61)
colnames(hfi)[pf]
```

Moreover, even those variables have been computed using other factors.
For instance we can observe pf_rol's composition as the following:
```{r}
colnames(hfi[5:8])
```


## 2. Descriptive Statistics (Niklas)

## 3. Principal Component Analysis (PCA) (Adriana)

## 4. Cluster Analysis (Niklas)

## 5. Multidimensional Scaling (Adriana)

## 6. Perceptual Maps (Guglielmo)

## 7. Possible Extension (Guglielmo) (NEW HEADLINE NEEDED)
